Estimating the dynamics of\\SPARQL query results over Wikidata\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}

### 2019-01-26 kgswc19.tex
Some further edits in Section 4

@@

@@

\paragraph{Feature Extractor:} Our features are based on the syntactic analysis of the queries; then some features are delivered to the Dynamics Model, which provides an indicator based on the probability of change for the constant predicates within the triple patterns of the query.

@@

Figure~\ref{fig:query} shows the complete list of evaluated features and shows the nominals. In the case of negation, we consider the statements MINUS, FILTER NOT EXISTS, and FILTER NOT BOUND and by the recursive path we understand those Property Paths that have patterns of the form $e*$ or $e+$ (see Definition~\ref{eq:pp}).

\begin{lstlisting}[captionpos=b, caption=Query example, label=fig:query,
basicstyle=\ttfamily,frame=single]
SELECT ?item
WHERE {
  ?item :instance_of :human .
  ?item :gender :female .
    { ?item :place_of_birth :Wales }
  UNION
    { ?item :place_of_birth ?pob .
      ?pob :located_in* :Wales }
  OPTIONAL { ?sitelink schema:about ?item .
             ?sitelink schema:inLanguage "cy" }
  FILTER (!BOUND(?sitelink))
}
LIMIT 100
\end{lstlisting}

\begin{center}
	\begin{tabular}{|l|ll|}    \hline
		1  & \# triples        & 7                             \\
		2  & \# vars           & 3                             \\
		3  & \# projected vars & 1                             \\
		4  & \# predicates     & 6                             \\
		5  & FILTER            & TRUE                          \\
		6  & LIMIT             & TRUE                          \\
		7  & UNION             & TRUE                          \\
		8  & SUBQUERY          & FALSE                         \\
		9  & *PATH             & TRUE                          \\
		10 & NEGATION          & TRUE                          \\
		11 & predicates        & {[}$p_1$,...,$p_k${]}    \\    \hline
	\end{tabular}
\end{center}

>>>>

\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{xspace}

\newcommand{\tickYes}{\checkmark}
\newcommand{\tickNo}{\textsl{\textsf{X}}}
\newcommand{\ntextnumero}{{\fontfamily{txr}\selectfont \textnumero}\xspace}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% SPARQL Listing style
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{listings}

\colorlet{punct}{red!60!black}
\definecolor{background}{HTML}{EEEEEE}
\definecolor{delim}{RGB}{120,20,40}
\definecolor{keyw}{RGB}{0,0,192}
\colorlet{numb}{magenta!60!black}

\lstdefinelanguage{sparql}{
	sensitive=false,
	extendedchars=true,
	literate={á}{{\'a}}1 {é}{{\'e}}1 {í}{{\'{\i}}}1 {ó}{{\'o}}1 {ú}{{\'u}}1
	{Á}{{\'A}}1 {É}{{\'E}}1 {Í}{{\'I}}1 {Ó}{{\'O}}1 {Ú}{{\'U}}1
	{ü}{{\"u}}1 {Ü}{{\"U}}1 {ñ}{{\~n}}1 {Ñ}{{\~N}}1 {¿}{{?``}}1 {¡}{{!``}}1
	{<}{{{\color{delim}<}}}{1}
	{>}{{{\color{delim}>}}}{1}
	{?}{{{\color{delim}?}}}{1}
	{*}{{{\color{delim}*}}}{1}
	{+}{{{\color{delim}+}}}{1}
	{/}{{{\color{delim}/}}}{1}
	{,}{{{\color{punct}{,}}}}{1}
	{;}{{{\color{punct}{;}}}}{1}
	{.}{{{\color{punct}{.}}}}{1}
	{:}{{{\color{punct}{:}}}}{1}
	{\{}{{{\color{delim}{\{}}}}{1} {\}}{{{\color{delim}{\}}}}}{1},
	morekeywords={ask,select,from,where,order,by,distinct,limit,offset,optional,union,filter,prefix,bound,desc,regex,str,group,not,exists,minus,service,certain,maybe}
}

\lstdefinestyle{sparqld}{
	basicstyle=\scriptsize\ttfamily,
	identifierstyle=\color{black},
	keywordstyle=\color{keyw}\bfseries,
	ndkeywordstyle=\color{greenCode}\bfseries,
	stringstyle=\color{ocherCode}\ttfamily,
	commentstyle=\color{darkgray}\ttfamily,
	language={sparql},
	tabsize=2,
	showtabs=false,
	showspaces=false,
	showstringspaces=false,
	extendedchars=true,
	escapechar=`,
	frame={single},
	breaklines=true,
	basewidth=0.5em,
	moredelim=[is][\color{magenta}]{~}{~},
	moredelim=**[is][\color{gray}]{£}{£},
	moredelim=**[is][\color{blue!50!black}]{$}{$},
	moredelim=**[is][\color{green!50!black}]{¬}{¬},
	xleftmargin=2ex,
	xrightmargin=1ex,
	aboveskip=1.5ex,
	belowskip=1.5ex
}

\makeatletter
\newcommand{\sqbox}{%
	\collectbox{%
		\@tempdima=\dimexpr\width-\totalheight\relax
		\ifdim\@tempdima<\z@
		\fbox{\hbox{\hspace{-.5\@tempdima}\BOXCONTENT\hspace{-.5\@tempdima}}}%
		\else
		\ht\collectedbox=\dimexpr\ht\collectedbox+.5\@tempdima\relax
		\dp\collectedbox=\dimexpr\dp\collectedbox+.5\@tempdima\relax
		\fbox{\BOXCONTENT}%
		\fi
	}%
}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% /SPARQL Listing style
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage{floatrow}
% Table float box with bottom caption, box width adjusted to content
%\newfloatcommand{capbtabbox}{table}[][\FBwidth]

@@

\paragraph{Feature Extractor:} Our initial set of features are based on analysis of the input SPARQL query. Figure~\ref{fig:query} shows the complete list of evaluated features and shows the nominals. In the case of negation, we consider the statements MINUS, FILTER NOT EXISTS, and FILTER NOT BOUND and by the recursive path we understand those Property Paths that have patterns of the form $e*$ or $e+$ (see Definition~\ref{eq:pp}).

@@

\begin{figure}[!t]
{\centering
	\begin{minipage}{0.46\textwidth}
	\begin{lstlisting}[style=sparqld]
SELECT ?item
WHERE {
  ?item :instance_of :human .
  ?item :gender :female .
    { ?item :place_of_birth :Wales }
  UNION
    { ?item :place_of_birth ?pob .
      ?pob :located_in* :Wales }
  OPTIONAL 
    { ?sitelink schema:about ?item .
      ?sitelink schema:inLanguage "cy" }
  FILTER (!BOUND(?sitelink))
}
LIMIT 100
	\end{lstlisting}
	\end{minipage}}
	\hfill
	\resizebox{0.46\textwidth}{!}{
	\begin{tabular}{rlr}    \toprule
		1  & \ntextnumero of triples        & 7			\\
		2  & \ntextnumero of variables      & 3			\\
		3  & \ntextnumero of projected vars & 1			\\
		4  & \ntextnumero of predicates     & 6			\\ \midrule
		5  & \texttt{FILTER}            	& \tickYes	\\
		6  & \texttt{LIMIT}             	& \tickYes	\\
		7  & \texttt{UNION}             	& \tickYes	\\
		8  & \textit{Sub-query}         	& \tickNo	\\
		9  & \textit{Recursive path}		& \tickYes	\\
		10 & \textit{Negation}         		& \tickYes	\\ \midrule
		11 & \textit{Predicates}			& $\{$ \texttt{:instance\_of}, ... $\}$     \\    \bottomrule
	\end{tabular}	
	}
\caption{Example query and the features extracted from it \label{fig:query}}	
\end{figure}
